
\documentclass[11pt,twoside,a4paper]{article}

\usepackage{amsmath,amsfonts,enumitem,mathrsfs,amsthm,amssymb,graphicx}
\usepackage[utf8]{inputenc}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{cor}{Corollary}

\theoremstyle{definition}
\newtheorem*{defn}{Definition}
\newtheorem{exmp}{Example}[section]


\title{The Classification Theorem for Compact Surfaces}
\author{Jishnu Kaiwar \\ [1cm]{\small Supervisor: Dr. Viveka Erlandsson} \\ Level 6, 20cp}

\begin{document}

\maketitle

\begin{abstract}
  The classification theorem for compact surfaces tells us that any
  compact surface is homeomorphic to one of a well known collection of
  surfaces connected sums of tori and projective planes, which we will define
  here). This is a remarkable result whose proof is easily accessible
  to readers without much background knowledge. In the first part of
  the project, we will start by introducing topological spaces,
  develop the theory till we can talk about ``continuity'' in these
  spaces. Finally we introduce the quotient topology which will give
  us a toolkit to complete the proof of the main theorem and give us many examples of surfaces

  The second part, comprising the proof of the main theorem, follows
  very intuitively given the preliminaries we develop whilst
  describing the quotient topology.
\end{abstract}

\newpage

\tableofcontents

\include{intro}
\include{prelims}
\include{proof}


\begin{thebibliography}{9}
\bibitem{munkres}  MUNKRES, J. R. (1974). \emph{Topology; a first course.} Englewood Cliffs, N.J., Prentice-Hall.
\bibitem{thom}  THOMASSEN, C. (1992). The Jordan-Sch\"onflies Theorem and the Classification of Surfaces. \emph{The American Mathematical Monthly}, 99(2), p.116.
\bibitem{mass}  MASSEY, W. (1996). \emph{Algebraic Topology.} New York: Springer.
\end{thebibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
